---
title: "Stat 333 - 2/23"
author: "Tess Steplyk"
date: "February 22, 2016"
output: html_document
---

###ISLR 4.7.6 
Suppose we collect data for a group of students in a statistics class with variables X1 = hours studied, X2 = undergrad GPA, and Y = receive an A. We fit a logistic regression and produce estimated co- efficients, βˆ0 = −6, βˆ1 = 0.05, βˆ2 = 1.
\
(a) Estimate the probability that a student who studies for 40 hours and has an undergrad GPA of 3.5 gets an A in the class.
\
Equation 4.2
```{r}
βo = -6
βl = 0.05
Xl = 40
X2 = 3.5
pX = (exp(βo + βl*Xl + X2))/(1 + exp(βo + βl*Xl + X2))
pX
```
***The probability is about 38%.***
\
(b) How many hours would the student in part (a) need to study to have a 50% chance of getting an A in the class?
\
***FIND XX. \
.5 = (exp(βo + βl*XX + X2))/(1 + exp(βo + βl*XX + X2)) \
1 = exp(βo + βl*XX + X2) \
XX = 2.5/.05 \
XX = 50***
\


###ISLR 4.7.7.
Suppose that we wish to predict whether a given stock will issue a dividend this year (“Yes” or “No”) based on X, last year’s percent profit. We examine a large number of companies and discover that the mean value of X for companies that issued a dividend was X ̄ = 10, while the mean for those that didn’t was X ̄ = 0. In addition, the variance of X for these two sets of companies was σˆ2 = 36. Finally, 80% of companies issued dividends. Predict the probability that a company will issue a dividend this year given that its percentage return was X = 4 last year.
\
Equation 4.12
It suffices to plug in the parameters and XX values in the equation for pk(x)pk(x). We get
so the probability that a company will issue a dividend this year given that its percentage return was X=4X=4 last year is 0.7520.752.

```{r}
pl = (0.8 * exp(-((1/72)*36))) / (0.8 * exp(-((1/72)*36)) + 0.2 * exp(-((1/72)*16)))
pl
```
***The probability a company will issue a dividend this year given its percentage return was X=4 last year is 0.752, or 75.2%.***
